Herring Case Study

### 2016-11-10 diversity/analysis/herring_case_study_v2.Rmd
herring case study

---
title: "Herring: case study 1"
author: ''
date: ''
output:
  pdf_document:
    fig_caption: yes
  html_document: default
  word_document: default
---

@@

## The Boom and Bust of Pacific Herring

@@

## Shifting demand: the graying of the palette

@@

Trends in revenue and landings in Prince William Sound fisheries largely mirror those of the rest of the state, with peaks in 1988 (Carlson 2005). Coincident with EVOS in 1989, harvest and prices  and declines through the 1990s.

In addition to being susceptible to changing demand, a related pressure on Alaskan sac roe fisheries has been external markets. Over the period 1994-2003 for example, roe from Alaska represented a small fraction of the total supplied to Japan, especially compared to roe from Canada and China (Carlson 2005).

## Changing Participation

Declines in herring abundnace, demand (and prices) and ex-vessel values in Prince William Sound have been well documented (Carlson 2005). Following peak

Participation

@@

```{r, echo=FALSE, fig.pos="placeHere", fig.cap = "Fishery landings for herring in Prince William Sound. L12E and L21E permits are combined for spawn on kelp, H01E and H34E are combined for food and bait, and G01E and G34E are combined for sac roe fisheries."}

@@

g2 = group_by(pws[pws$year<=1998,], Fishery, year) %>% 
  summarize(w = length(unique(p_holder))) %>% 
  ggplot(aes(year, w, group=Fishery, color=Fishery)) + geom_line() + 
  xlab("Year") + ylab("Active permits")

gridExtra::grid.arrange(g1, g2)
```

## Changing Strategies

```{r, eval=FALSE}
cfec <- feather::read_feather("../../portfolio/data-generated/cfec.feather")
cfec$year <- as.numeric(cfec$year)

this.permit = "G 01E"
# S01E S03T - effective permit diversity

simp.div = function(x) {
  1/sum((x/sum(x))^2)
}
```

```{r, eval=FALSE}
g1 = group_by(cfec, p_holder) %>% 
  mutate(pws = ifelse(length(which(p_fshy==c(this.permit)))>0, 1, 0)) %>% 
  filter(pws>0) %>%
  select(-pws) %>% 
  group_by(p_holder, year, p_fshy) %>% 
  summarize(g = sum(g_earn)) %>%
  group_by(p_holder, year) %>% 
  summarize(div = simp.div(g)) %>% 
  group_by(year) %>% 
  summarize(mean = mean(div), lower=quantile(div,0.025), upper=quantile(div,0.975)) %>% 
  ggplot(aes(year, mean)) + geom_line() + ggtitle(this.permit) + ylab("Mean diversity")

# Identify dominant strategies associated with this permit
g2 = group_by(cfec, p_holder) %>% 
  mutate(pws = ifelse(length(which(p_fshy==c(this.permit)))>0, 1, 0)) %>% 
  filter(pws>0) %>%
  select(-pws) %>% 
  group_by(p_holder, year, p_fshy) %>% 
  summarize(g = sum(g_earn)) %>%
  group_by(p_holder, year) %>% 
  mutate(totRev = sum(g)) %>% 
  group_by(p_holder, year, p_fshy) %>% 
  mutate(g = g / totRev) %>% 
  select(-totRev)

# use reshape 2 to cast this to wide format
aql <- melt(g2, id.vars = c("year", "p_fshy", "p_holder"))
aqw <- dcast(aql, year + p_holder ~ p_fshy)
# replace NAs w/0s
for(i in 1:ncol(aqw)) {
  aqw[which(is.na(aqw[,i])),i] = 0
}

indx = sample(seq(1,nrow(aqw)),size=500,replace=F)
sub = aqw[indx,-c(1,2)]

sub = aqw[,-c(1,2)]
coarse_df = aqw[,-c(1,2)]
coarse_df$S = apply(coarse_df[,which(substr(names(sub),1,1)=="S")],1,sum)
coarse_df$B = apply(coarse_df[,which(substr(names(sub),1,1)=="B")],1,sum)
coarse_df$GHL = apply(coarse_df[,which(substr(names(sub),1,1)%in%c("G","H","L"))],1,sum)
coarse_df$KTD = apply(coarse_df[,which(substr(names(sub),1,1)%in%c("K","T","D"))],1,sum)
coarse_df$M = apply(coarse_df[,which(substr(names(sub),1,1)%in%c("M"))],1,sum)
coarse_df$P = apply(coarse_df[,which(substr(names(sub),1,1)%in%c("P"))],1,sum)
coarse_df = coarse_df[,c("S","B","GHL","KTD","M","P")]

clust = fpc::pamk(coarse_df[sample(seq(1,nrow(aqw)),size=500,replace=F),])# cluster based on subset to pick medoids
medoids = clust$pamobject$medoids
clust = pam(coarse_df, k = clust$nc) # cluster full dataset

names_medoids = ""
for(i in 1:nrow(medoids)) {
  names_medoids[i] = paste0(colnames(medoids)[which(medoids[i,] > 0.05)], collapse=":")
}

aqw$clusters = clust$clustering
totals = group_by(aqw, clusters, year) %>% 
  summarize(n = n())
totals$group = names_medoids[totals$clusters]

ggplot(totals, aes(year, n, group=group, color=group)) + geom_line() + geom_point()

ggplot(totals, aes(year, n)) + geom_area(aes(fill = group), position = "stack", colour = 1)

```

```{r,eval=FALSE}
cfec <- feather::read_feather("../../portfolio/data-generated/cfec.feather")
cfec$year <- as.numeric(cfec$year)

# grab people who ever held prince william sound herring permit
# calculate unique permits by person-year
pws.G34 = group_by(cfec, p_holder) %>% 
  mutate(pws = ifelse(length(which(p_fshy==c("G 34E")))>0, 1, 0)) %>% 
  filter(pws>0) %>%
  select(-pws) %>% 
  group_by(p_holder, year) %>% 
  summarize(n = length(unique(p_fshy))) %>%
  group_by(year) %>% 
  summarize(p1 = length(which(n>1)) / length(which(n > 0)))

pws.G01 = group_by(cfec, p_holder) %>% 
  mutate(pws = ifelse(length(which(p_fshy==c("G 01E")))>0, 1, 0)) %>% 
  filter(pws>0) %>%
  select(-pws) %>% 
  group_by(p_holder, year) %>% 
  summarize(n = length(unique(p_fshy))) %>%
  group_by(year) %>% 
  summarize(p1 = length(which(n>1)) / length(which(n > 0)))

pws.G34$permit="g34"
pws.G01$permit="g01"
pws = rbind(pws.G34, pws.G01)

ggplot(pws, aes(year, p1, group=permit, color=permit)) + geom_line() + 
  ylab("Proportion of people landing fish under > 1 permit") + xlab("Year")

```

```{r eval=FALSE}
# grab people who ever held prince william sound herring permit
# calculate unique permits by person-year
pws.G34 = group_by(cfec, p_holder) %>% 
  mutate(pws = ifelse(length(which(p_fshy==c("G 34E")))>0, 1, 0)) %>% 
  filter(pws>0) %>%
  select(-pws) %>% 
  group_by(p_holder, year) %>% 
  summarize(n = length(unique(p_fshy))) %>%
  group_by(year) %>% 
  summarize(p1 = length(which(n>1)) / length(which(n > 0)))

pws.G01 = group_by(cfec, p_holder) %>% 
  mutate(pws = ifelse(length(which(p_fshy==c("G 01E")))>0, 1, 0)) %>% 
  filter(pws>0) %>%
  select(-pws) %>% 
  group_by(p_holder, year) %>% 
  summarize(n = length(unique(p_fshy))) %>%
  group_by(year) %>% 
  summarize(p1 = length(which(n>1)) / length(which(n > 0)))

pws.G34$permit="g34"
pws.G01$permit="g01"
pws = rbind(pws.G34, pws.G01)

ggplot(pws, aes(year, p1, group=permit, color=permit)) + geom_line() + 
  ylab("Proportion of people landing fish under > 1 permit") + xlab("Year")

```

```{r, echo = FALSE, warning = FALSE, message = FALSE, results = "hide", eval=FALSE}
cfec <- feather::read_feather("../../portfolio/data-generated/cfec.feather")
cfec$year <- as.numeric(cfec$year)

# grab people who ever held prince william sound herring permit
# calculate unique permits by person-year
pws.G34 = group_by(cfec, p_holder) %>% 
  mutate(pws = ifelse(length(which(p_fshy==c("G 34E")))>0, 1, 0)) %>% 
  filter(pws>0) %>%
  select(-pws) %>% 
  group_by(p_holder, year) %>% 
  summarize(n = length(unique(p_fshy))) %>%
  group_by(year) %>% 
  summarize(p1 = length(which(n>1)) / length(which(n > 0)))

pws.G01 = group_by(cfec, p_holder) %>% 
  mutate(pws = ifelse(length(which(p_fshy==c("G 01E")))>0, 1, 0)) %>% 
  filter(pws>0) %>%
  select(-pws) %>% 
  group_by(p_holder, year) %>% 
  summarize(n = length(unique(p_fshy))) %>%
  group_by(year) %>% 
  summarize(p1 = length(which(n>1)) / length(which(n > 0)))

pws.G34$permit="g34"
pws.G01$permit="g01"
pws = rbind(pws.G34, pws.G01)

ggplot(pws, aes(year, p1, group=permit, color=permit)) + geom_line() + 
  ylab("Proportion of people landing fish under > 1 permit") + xlab("Year")

herring = cfec[substr(cfec$p_fshy,1,1) %in% c("L","G","H"),]

herring_summary = group_by(herring, p_fshy, year) %>%
  summarize(g = sum(g_earn), p = mean(g_price,na.rm=T), lbs = sum(g_pounds), people = length(unique(p_holder))) %>%
  group_by(p_fshy) %>%
  mutate(n = length(unique(year)), maxpeople = max(people,na.rm=T)) %>%
  filter(n > 9) %>%
  filter(maxpeople > 25)

quitters = group_by(herring, p_holder, p_fshy) %>%
  summarize(lastYear = max(year))

group_by(quitters[quitters$lastYear<2014,],lastYear,p_fshy) %>% 
  summarize(n = n()) %>% 
  ggplot(aes(lastYear,n)) + geom_line() + facet_wrap(~p_fshy, scale="free_y") + xlab("Year") + ylab("People leaving the fishery (last year fished)")

# look at what these people did after leaving
cfec_quitters = cfec[which(cfec$p_holder%in%unique(quitters$p_holder)),]
cfec_quitters = left_join(cfec_quitters, quitters[,c("p_holder","lastYear")])

fates = as.data.frame(group_by(cfec_quitters, p_holder, p_fshy) %>% 
  summarize(n = length(which(year > lastYear[1]))) %>% 
  group_by(p_fshy) %>% 
  summarize(n0 = length(which(n==0)), nOther = length(which(n>0))))

trackQuitters = function(permit) {
cfec2= cfec_quitters[which(cfec_quitters$p_holder %in% unique(quitters[quitters$p_fshy==permit,]$p_holder)),]
cfec2 = cfec2[which(cfec2$year > cfec2$lastYear),]
cfec2$SPEC = substr(cfec2$p_fshy,1,1)
totalpermit = group_by(cfec2,SPEC) %>% 
  summarize(g = sum(g_earn))
totalpermit$g = totalpermit$g/sum(totalpermit$g)
return(as.data.frame(arrange(totalpermit,-g)))

}

#trackQuitters("L 21C")
```

## Overview

First, we can pull out the herring permits by all areas. We'll restrict the analysis to (1) fisheries with 10 or more years of data, and (2) fisheries that at their max had at least 25 people landing catch in a year. We see that people generally decline to 0, except fisheries in SE (G01A, G34A), and spawn on kelp fisheries (L21). Despite catch being flat or maybe increasing, we see prices go down or flat for nearly all permits (and spawn on kelp L21 permits have significantly higher price / lb compared to roe fisheries).

In terms of the diversification of these permit holders, some questions to ask might be (1) if people stopped fishing herring, did they stop fishing entirely, or fish something else? (2) for the people that stayed, how many of them are diversified (other permits) and how much of their revenue comes from other permits?

For a lot of permits, we see decreases in the proportion of people fishing only herring. In terms of spatial diversity, for several of the fisheries that haven't crashed, we see increases in diversity: G01M/T, G34N, L21. In terms of diversifying to other types of herring permits,

```{r, eval=FALSE, echo = FALSE, warning = FALSE, message = FALSE, results = "hide", fig.pos="placeHere", fig.cap = "Herring permit holders by year."}
  ggplot(herring_summary, aes(year, people)) + geom_line() + facet_wrap(~ p_fshy, scale="free_y") + xlab("Year") + ylab("People")
```

```{r, eval=FALSE, echo = FALSE, warning = FALSE, message = FALSE, results = "hide", fig.pos="placeHere", fig.cap = "Trends in landed catch by permit.", fig.height=3.5}
  ggplot(herring_summary, aes(year, log(lbs))) + geom_line() + 
  facet_wrap(~ p_fshy) + ylab("Ln (tons)") + xlab("Year")
```

```{r, eval=FALSE,echo = FALSE, warning = FALSE, message = FALSE, results = "hide", fig.pos="placeHere", fig.cap = "Trends in prices by permit.", fig.height=3.5}
  ggplot(herring_summary, aes(year, p)) + geom_line() + 
  facet_wrap(~ p_fshy, scale="free_y") + ylab("Price (Dollars / lb.)") + xlab("Year")
```

```{r, eval=FALSE,echo = FALSE, warning = FALSE, message = FALSE, results = "hide", fig.pos="placeHere", fig.cap = "Scatterplots of prices versus revenue by permit."}
  ggplot(herring_summary, aes(p, log(lbs/2000), colour = year)) + geom_point() + facet_wrap(~ p_fshy, scale="free") + ylab("Ln (tons)") + xlab("Price (dollars / lb.)")

```

## How much of your revenue is driven by non-herring?

```{r,eval=FALSE, echo = FALSE}

herring_permits = unique(herring_summary$p_fshy)

# the complicated thing here is that people may hold multiple herring permits, 
# multiple permits, no permits. So for each permit type we need to calculate 
# each of these groups independently. Some people may be counted multiple times.

# next calculate strategies for diversification
calc_permit_summary = function(this_permit) {
people_year = group_by(cfec[cfec$p_fshy == this_permit,], p_holder, year) %>% 
  summarize(g = sum(g_earn)) %>% 
  filter(g > 0) %>% 
  select(-g)
# grab all records pertaining to these people-year combinations
herring = left_join(people_year, cfec)

# aggregate fish tickets to permits
herring = group_by(herring, p_holder, p_fshy, year) %>% 
  summarize(g_earn = sum(g_earn))

permit_summary = group_by(herring, p_holder, year) %>% 
  summarize(npermit = length(unique(p_fshy)),
    npermit_herring = length(which(p_fshy %in% herring_permits)),
    npermit_g34 = length(which(p_fshy %in% herring_permits & substr(p_fshy,1,4)=="G 34")),
    npermit_g01 = length(which(p_fshy %in% herring_permits & substr(p_fshy,1,4)=="G 01")),
    npermit_l12 = length(which(p_fshy %in% herring_permits & substr(p_fshy,1,4)=="L 12")),
    npermit_l21 = length(which(p_fshy %in% herring_permits & substr(p_fshy,1,4)=="L 21")),
    nmult = ceiling(npermit_g34/100) + ceiling(npermit_g01/100) + ceiling(npermit_l21/100) + ceiling(npermit_l12/100)) %>% 
    as.data.frame() %>%
    group_by(year) %>% 
  summarize(only_herring_single = length(which(npermit==1 & npermit_herring == 1)),
    only_herring_multiple = length(which(npermit>1 & npermit_herring >1 & npermit == npermit_herring)), herring_combo = length(which(npermit > 1 & npermit_herring >= 1 & npermit > npermit_herring)), g34 = length(which(npermit_g34 > 1)), g01 = length(which(npermit_g01 > 1)),
    l12 = length(which(npermit_l12 > 1)), l21 = length(which(npermit_l21 > 1)), 
    multiple_fisheries = length(which(nmult > 1)))

permit_summary$p_fshy = this_permit
return(permit_summary)
}

herring_summary = calc_permit_summary(herring_permits[1])
for(i in 2:length(herring_permits)) {
  herring_summary = rbind(herring_summary, calc_permit_summary(herring_permits[i]))
}

```

```{r, eval=FALSE,echo = FALSE, warning = FALSE, message = FALSE, results = "hide", fig.pos="placeHere", fig.cap = "Proportion of permit holders specializing on herring.", fig.height=3.5}
ggplot(herring_summary, aes(year, (only_herring_single+only_herring_multiple)/(only_herring_single+only_herring_multiple+herring_combo))) + facet_wrap(~p_fshy) + geom_line() + xlab("Year") + ylab("Proportion of people only fishing herring")
```

```{r, eval=FALSE,echo = FALSE, warning = FALSE, message = FALSE, results = "hide", fig.pos="placeHere", fig.cap = "Proportion of permit holders diversifying spatially.", fig.height=3.5}
p1 = ggplot(herring_summary, aes(year, (g34)/(only_herring_single+only_herring_multiple+herring_combo), group = p_fshy, colour = p_fshy)) + geom_line() + xlab("Year") + ylab("Proportion of people diversifying spatially")

p2 = ggplot(herring_summary, aes(year, (g01)/(only_herring_single+only_herring_multiple+herring_combo), group = p_fshy, colour = p_fshy)) + geom_line() + xlab("Year") + ylab("Proportion of people diversifying spatially")

p3 = ggplot(herring_summary, aes(year, (l21)/(only_herring_single+only_herring_multiple+herring_combo), group = p_fshy, colour = p_fshy)) + geom_line() + xlab("Year") + ylab("Proportion of people diversifying spatially")

p4 = ggplot(herring_summary, aes(year, (l12)/(only_herring_single+only_herring_multiple+herring_combo), group = p_fshy, colour = p_fshy)) + geom_line() + xlab("Year") + ylab("Proportion of people diversifying spatially")

gridExtra::grid.arrange(p1, p2, p3, p4)
```

```{r, eval=FALSE,echo = FALSE, warning = FALSE, message = FALSE, results = "hide", fig.pos="placeHere", fig.cap = "Proportion of people diversifying w/diff herring permits.", fig.height=3.5}
ggplot(herring_summary, aes(year, (multiple_fisheries)/(only_herring_single+only_herring_multiple+herring_combo))) + facet_wrap(~p_fshy) + geom_line() + xlab("Year") + ylab("Proportion of people diversifying w/diff herring permits")
```

```{r, eval=FALSE,echo = FALSE, warning = FALSE, message = FALSE, results = "hide", fig.pos="placeHere", fig.cap = "Ratio of effective diversity to total number of permit holders (as measure of evenness). Values near 1 indicate everyone's revenue is similar.", fig.height=3.5}
# We can also calculate how herring catch is distributed across fisherment
simp.div = function(x) {1/sum((x/sum(x))^2)}

group_by(cfec[cfec$p_fshy%in%herring_permits,], year, p_holder, p_fshy) %>% 
  summarize(g = sum(g_earn)) %>% 
  group_by(year, p_fshy) %>% 
  summarize(div = simp.div(g), n = length(unique(p_holder))) %>% 
  ggplot(aes(year, div/n)) + facet_wrap(~p_fshy, scale="free_y") + geom_line() + xlab("Year") + ylab("Diversity across people")
```

```{r, eval=FALSE,echo = FALSE, warning = FALSE, message = FALSE, results = "hide", fig.pos="placeHere", fig.cap = "Fraction of revenue made up of herring (median, of per-person fractions).",eval=FALSE}

herring_permits = unique(herring_summary$p_fshy)

herring_people = group_by(cfec, p_fshy, year, p_holder) %>%
  summarize(g = sum(g_earn),
  herring_landing = ifelse(length(which(p_fshy %in% herring_permits)) == 0, 0, 1)) %>% 
  data.frame() %>% 
  group_by(p_holder, year) %>% 
  mutate(herring_personyear = ifelse(length(which(p_fshy %in% herring_permits)) == 0, 0, 1),
  g_herring = sum(g[herring_landing==1]), g_other = sum(g[herring_landing==0]),
  p_fshy = p_fshy[which(p_fshy %in% herring_permits)][1]) %>% 
  group_by(p_holder) %>% 
  mutate(ever_herring = sum(herring_personyear)) %>%   
  filter(ever_herring > 0) %>% 
  select(-g, -herring_landing, -herring_personyear, -ever_herring)

herring_people$p_herring = herring_people$g_herring / (herring_people$g_herring + herring_people$g_other)
group_by(herring_people[herring_people$p_fshy %in% herring_permits, ], p_fshy, year) %>% 
  summarize(g = median(p_herring)) %>% 
  ggplot(aes(year, g)) + geom_line() + facet_wrap(~p_fshy) + ylab("Median % herring (as fraction of total revenue)") + xlab("Year")
```

## References

@@

@@

>>>>

---
title: 'Herring Case Study'
author: ''
date: ''
output:
  word_document: default
  pdf_document:
    fig_caption: yes
  html_document: default
---

@@

### The Boom and Bust of Pacific Herring

@@

### Shifting demand: the graying of the palate

@@

Trends in revenue, biomass, and landings in Prince William Sound fisheries largely mirror those of the rest of the state, with peaks in 1988 (Pearson et al. 2012; Carlson 2005). Coincident with EVOS in 1989, prices and harvest declined through much of the 1990s, with limited fisheries openings, and fisheries have remained closed since 1999 because estimated biomass has remained lower than that required for reopening the fishery (20000 short tons). Though not affected by EVOS, harvest of herring roe in other Alaska fisheries have also declined since the late 1980s; these include fisheries in Norton Sound, Bristol Bay, and Togiak (Woodby et al. 2005). Perhaps the exception to this declining trend is roe harvest is in Sitka Sound (Thynes et al. 2016;  Woodby et al. 2005). The timing of herring spawn and harvest typically follows a latitudinal gradient from south to north (Carlson 2005), so more southern fisheries and processors may have an advantage on fisheries in other areas.

In addition to being susceptible to changing demand, a related pressure on Alaskan sac roe fisheries has been external markets. Over the period 1994-2003 for example, roe from Alaska represented a small fraction of the total supplied to Japan, especially compared to roe from Canada and China (Carlson 2005). Though roe from Pacific herring are considered to be of the highest quality, there may also be competition from other roe fisheries offering substitutes (e.g. Atlantic herring). Finally, because the salted roe is preserved, supply and demand may be affected by surplus in previous years (Carlson 2005).

@@

```{r, echo=FALSE, fig.pos="placeHere", fig.cap = "Active permits for herring in Prince William Sound. L12E and L21E permits are combined for spawn on kelp, H01E and H34E are combined for food and bait, and G01E and G34E are combined for sac roe fisheries."}

@@

g2 = group_by(pws[pws$year<=1998,], Fishery, year) %>% 
  summarize(w = length(unique(p_holder))) %>% 
  ggplot(aes(year, w, group=Fishery, color=Fishery)) + geom_line() + 
  xlab("Year") + ylab("Active permits")
print(g2)
#gridExtra::grid.arrange(g1, g2)
```

### Changing Participation

Declines in herring abundance, demand (and prices) and ex-vessel values in Prince William Sound have been well documented (Carlson 2005). Following peak biomass in the late 1980s, the number of active permit holders declined significantly until the fisheries closed in the late 1990s. Permits for herring roe fisherines in Prince William Sound may be unique to other case studies in that although the fishery remains closed, the permits continue to hold value (and are bought and sold). Because of the closure, permit values have declined substantially to approximately 10% of the value 20 years ago (from 127000 dollars in February 1997 or 327000 in 2016 dollars to 31000 in February 2016) (CFEC Permit Value Report 2016).

### Changing diversification: the salmon safety net

With herring fisheries in PWS closed, permit holders had two options: stop fishing entirely, or fish other non-herring permits that remained open. The fate of herring permit holders in PWS was evaluated using the CFEC database, 1985-2014. We defined this 'population' of permit holders as being anyone who ever held one of the six herring permits in PWS. In each year, we summarized the number of active fishers (on any permit, including non-herring fisheries). As a proxy for individual permit holders leaving fishing entirely, we also calculated the last year that each permit holder was active (again, on any permit). These summaries show a long-term decline in the number of individuals in this population (an expected result, as many of the herring permit holders had herring contribute a large fraction of their revenue). We also see fishery closures as catalysts for people stopping participation in the fishery.

```{r, echo=FALSE, fig.pos="placeHere", fig.cap="Histogram of last year individual permit holders holding Prince William Sound herring permits were active in any fishery throughout the state. The last years that herring fishing was permitted in Prince William Sound was 1997-1998."}
pws = dplyr::filter(cfec, p_fshy%in%c("G 01E", "G 34E", "H 01E", "H 34E", "L 12E", "L 21E"))
pws2 = cfec[which(cfec$p_holder%in%unique(pws$p_holder)),]

g = group_by(pws2, p_holder) %>% 
  summarize(maxY = max(year))

g2 = group_by(pws2, year) %>% 
  summarize(n = length(unique(p_holder)))
par(mfrow=c(2,1), mgp=c(2,1,0),mai=c(0.7,0.7,0.1,0.2))
plot(g2$year, g2$n, type="b", xlab="Year", ylab="Active permit holders", lwd=2, ylim=c(0,600))
plot(table(g$maxY)[1:29], xlab="Last year fished", ylab="Permit holders")
```

For the population of fishers who ever held a PWS herring permit, we summarized the revenue by year (revenue summed in dollars, using deflation corrected earnings). Individual permits were grouped to coarse taxonomic levels by the first character of their CFEC permit code (e.g. all 'G', 'H', and 'L' permits combined as herring, all 'S' permits combined for salmon, all 'B' permits combined for halibut; Figure XX). These aggregate groups indicate that as expected, herring make up a smaller fraction of the total revenue over time (particularly following fisheries closures). Herring continue to make up a small fraction of revenue for this population, because several fishers continue to be active in other areas throughout the state. Few of these individuals remain herring specialists however; a more common strategy is for fishers to actively fish both herring and salmon (Figure XX). For the majority of individuals in this population, the CFEC data indicate that the majority of active fishers abandonded herring entirely. The most common non-herring strategy for these fishers was to target salmon exclusively. A smaller number of fishers became specialized on halibut, or targeted both salmon and halibut in the same year. Unlike halibut, groundfish, or crab fisheries, where the ability to participate is more restrictive (permit holders must buy quota), limited entry salmon fisheries in the state may provide a safety net for fishers affected by the decline or closure of other fisheries in the state.

### References

@@

Pearson, W.H., R.B. Deriso, R.A. Elston, S.E. Hook, K.R. Parker, and J.W. Anderson. 2012. Hypotheses concerning the decline and poor recovery of Pacific herring in Prince William Sound, Alaska. Reviews in Fish Biology and Fisheries, 22(1): 95–135.

@@

Thynes, T., D. Gordon, D. Harris, and S. Walker. 2016. 2016 Southeast Alaska Sac Roe Herring Fishery Management Plan. Regional Information Report 1J16-02. ADFG, Division of Commercial Fisheries.
